---
title: "Machine Learning assignment"
author: "JayR"
date: "20/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synopsis

The aim of this project is to create a model that uses data to predict outcomes. The data being analysed is from accelerometers on the belt, forearm, arm and dumbell of 6 paritipants. The participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

Load required libraries
```{r, results = FALSE}
library(tidyverse)
library(caret)
library(skimr)
library(rattle)
library(gbm)
```

## Data analysis

Import data
```{r, results = FALSE}
training <- read_csv("./data/pml-training.csv")
testing <- read_csv("./data/pml-testing.csv")
```
```{r}
dim(training)
dim(testing)
```

View data
```{r}
head(training)
```

Look at how complete the columns are
```{r, results = FALSE}
skim(training)
```

There are several columns that are less than 3% complete - remove these columns from the training and testing datasets. Also remove any non-numeric variables (except `classe`)
```{r}
training_ed <- training %>%
  select_if(~ !any(is.na(.))) %>%
  select(-c(X1:num_window))

testing_ed <- testing %>%
  select_if(~ !any(is.na(.))) %>%
  select(-c(X1:num_window))
```

Set `classe` as a factor
```{r}
training_ed$classe <- as.factor(training_ed$classe)
```

## Data partitioning

Split the training data into training and testing datasets
```{r}
set.seed(100)
inTrain <- createDataPartition(training_ed$classe, p = 0.6, list = FALSE)

traindata <- training_ed[inTrain,]
testdata <- training_ed[-inTrain,]
dim(traindata)
dim(testdata)
```

## Use cross validation to construct models

# Decision Tree Model

```{r, results = FALSE}
modFitDecT <- train(classe ~., data = traindata, method = "rpart")
```
```{r}
fancyRpartPlot(modFitDecT$finalModel)
```



```{r}
predDecT <- predict(modFitDecT, testdata)
confDecT <- confusionMatrix(predDecT, testdata$classe)
confDecT
```

The Decision Tree model predicts with accuracy of `r confDecT$overall[["Accuracy"]]`.

```{r}
table(predDecT, testdata$classe)
```

```{r}
plot(confDecT$table, col = confDecT$byClass, main = "Decision Tree Predictions")
```

# Gradient Boosting Model

```{r, results = FALSE, cache = TRUE}
set.seed(25621)
modFitGBM <- train(classe ~., data = traindata, method = "gbm", verbose = FALSE)
```

```{r}
predGBM <- predict(modFitGBM, testdata)
confGBM <- confusionMatrix(predGBM, testdata$classe)
confGBM
```

The Gradient Boosting model predicts with accuracy of `r # confGBM$overall[["Accuracy"]]`.

```{r}
table(predGBM, testdata$classe)
```

```{r}
plot(confGBM$table, col = confGBM$byClass, main = "Gradient Boosting Predictions")
```


# Linear Discriminant Analysis model

```{r, results = FALSE}
modFitLDA <- train(classe ~., data = traindata, method = "lda")
```

```{r}
predLDA <- predict(modFitLDA, testdata)
confLDA <- confusionMatrix(predLDA, testdata$classe)
confLDA
```

The Linear Discriminant Analysis model predicts with accuracy of `r confLDA$overall[["Accuracy"]]`.

```{r}
table(predLDA, testdata$classe)
```

```{r}
plot(confLDA$table, col = confLDA$byClass, main = "Linear Discriminant Analysis Predictions")
```

# Random Forest Model

```{r, results = FALSE, cache = TRUE}
modFitRF <- train(classe ~., data = traindata, method = "rf", ntree = 100)
```

```{r, cache = TRUE}
predRF <- predict(modFitRF, testdata)
confRF <- confusionMatrix(predRF, testdata$classe)
confRF
```

The Random Forest model predicts with accuracy of `r confRF$overall[["Accuracy"]]`.

```{r}
table(predRF, testdata$classe)
```

```{r}
plot(confRF$table, col = confRF$byClass, main = "Random Forest Predictions")
```

## Final model chosen

The Random Forest model has produced the highest accuracy so is likely to be the best model to chose.

# Final prediction
```{r}
Final_pred <- predict(modFitRF, testing_ed)
Final_pred
```
